---
title: "WDI: ICT"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Last updated on: `r Sys.Date()`*

Cellphone and internet adoptions rates for all countries, 1960 on. 

- The WDI indicators used for this are ... 
- The data were changed to conform as much as possible to the Gleditsch & Ward state list. 
- The data are lagged 1 year so that there are 2019 values
- All missing values for series that were not completely missing for a country have been imputed. The strategy was:
  + ...

```{r}
library(ggplot2)
library(dplyr)

df <- read.csv("output/wdi-ict.csv")

glimpse(df)

ggplot(df, aes(x = year, y = internet_users_pct, group = gwcode)) +
  geom_line(alpha = 0.5) +
  theme_light()

# Scaled version that is adjusted for annual mean and sd
ggplot(df, aes(x = year, y = cellphones_per100, group = gwcode)) +
  geom_line(alpha = 0.5) +
  theme_light()
```

## Data cleaning

See [clean-data.md](clean-data.md) for results of the data cleaning script. 
