---
title: "Combine data into states.rds"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages({
  library(tidyverse)
  library(states)
  library(yaml)
})

# shortcut for plotting missing values when no date column is available
plotmiss <- function(x) {
  plot_missing(x, names(x), partial = "any")
}
```

## Pieces

### Master statelist

```{r}
states <- state_panel(1950, 2019, partial = "any")
```

### P&T coups

```{r}
ptcoups <- read_csv("input/ptcoups.csv")

glimpse(ptcoups)

plotmiss(ptcoups)

states <- left_join(states, ptcoups, by = c("gwcode", "year"))
```

### Make lead DV versions

```{r}
dv_vars <- ptcoups %>%
  select(gwcode, year, pt_attempt, pt_coup, pt_failed) 
lead1 <- dv_vars %>%
  mutate(year = year - 1) %>%
  rename(pt_attempt_lead1 = pt_attempt, pt_coup_lead1 = pt_coup, 
         pt_failed_lead1 = pt_failed)
lead2 <- dv_vars %>%
  mutate(year = year - 2) %>%
  rename(pt_attempt_lead2 = pt_attempt, pt_coup_lead2 = pt_coup, 
         pt_failed_lead2 = pt_failed)
dv <- full_join(lead1, lead2)
```

Cuba had an attempt in 1952, check data are correct for this:

```{r}
cuba <- dv %>%
  filter(gwcode==40 & year < 1953) %>%
  arrange(year) %>%
  select(gwcode, year, pt_coup_lead1, pt_coup_lead2)

cuba

stopifnot(cuba$pt_coup_lead1[cuba$year==1951] == 1,
          cuba$pt_coup_lead2[cuba$year==1950] == 1)
```

#### Impute for last 2 years in state existence

Several states drop out during the data period. Set the DV vars to 0 in those instances.

```{r}
last2 <- function(x, yy, ly) {
  ifelse(is.na(x) & yy >= ly - 1, 0L, x)
}

dv_full <- left_join(states, dv) %>%
  select(gwcode, year, contains("lead")) %>%
  group_by(gwcode) %>%
  mutate(last_year = max(year),
         # states with last year in 2019 still exist
         last_year = ifelse(last_year==max(states$year), 9999L, last_year)) %>%
  mutate_at(.vars = vars(contains("lead")), ~last2(., year, last_year)) %>%
  select(-last_year)
```

```{r}
states <- left_join(states, dv_full)
```

## G&W state age

Years since independence

```{r}
age <- read_csv("input/gwstate-age.csv") %>%
  mutate(ln_state_age = log(state_age)) %>%
  select(-state_age)

states <- left_join(states, age, by = c("gwcode", "year"))
```

## EPR

Ethnic Power Relations data on ethnic groups in countries.

```{r}
epr <- read_csv("input/epr.csv") %>%
  setNames(c("gwcode", "year", paste0("epr_", names(.)[3:ncol(.)])))

glimpse(epr)
```

Add year-to-year difference of these variables. Changes could be significant. 

```{r, eval=FALSE}
TODO
```


EPR only ranges to 2017 right now. Lag and carry-back, or don't lag but do carry-forward from 2017 on?

```{r}
ggplot(epr, aes(x = year, y = epr_excluded_group_pop, group = gwcode)) +
  geom_line()
```

There are quite frequent changes, as the plot above shows. Since I care more about recent cases, and since it seems fair to assume that the power structures pre-independence were similar to those at independence, lag and carry back for first two years in state's existence, for states that enter after 1950. 

```{r}
epr_lagged <- epr %>%
  mutate(year = year + 2) %>%
  arrange(gwcode, year) %>%
  group_by(gwcode) %>%
  tidyr::fill(-gwcode, -year, .direction = "up")
```

```{r}
states <- left_join(states, epr_lagged, by = c("gwcode", "year"))
```

## REIGN data

```{r}
reign <- read_csv("input/reign-cy.csv") %>%
  setNames(., c("gwcode", "year", paste0("reign_", names(.)[3:ncol(.)])))

glimpse(reign)

# Simplify the government coding
table(reign$reign_government) %>% sort() %>% rev()
```
```{r}
check <- left_join(states, reign)
check %>% 
  mutate(pt_attempt_lead1 = as.integer(as.character(pt_attempt_lead1)),
         pt_coup_lead1 = as.integer(as.character(pt_coup_lead1))) %>%
  group_by(reign_government) %>% 
  summarize(n = n(), 
            attempts = sum(pt_attempt_lead1, na.rm = TRUE), 
            attempt_rate = mean(pt_attempt_lead1, na.rm = TRUE), 
            coup = sum(pt_coup_lead1, na.rm = TRUE), 
            coup_rate = mean(pt_coup_lead1, na.rm = TRUE)) %>% 
  arrange(desc(attempt_rate))
```

```{r}
reign <- reign %>%
  mutate(
    reign_gov_pres = as.integer(reign_government=="Presidential Democracy"),
    reign_gov_parl = as.integer(reign_government=="Parliamentary Democracy"),
    reign_gov_personal = as.integer(reign_government %in% c("Personal Dictatorship", "Monarchy")),
    reign_gov_party = as.integer(reign_government %in% c("Dominant Party", "Party-Personal", "Party-Personal-Military Hybrid", "Party-Military")),
    reign_gov_provisional = as.integer(str_detect(reign_government, "Provisional")),
    reign_gov_military = as.integer(reign_government %in% c("Indirect Military", "Military", "Warlordism", "Military-Personal"))
  ) %>%
  select(-reign_government)

states <- left_join(states, reign)
```

## V-Dem

```{r}
vdem <- read_csv("input/v-dem.csv",
                 col_types = cols(
                   .default = col_double(),
                   country_name = col_character(),
                   year = col_integer()
                 )) %>%
  select(-country_name) %>%
  filter(!is.na(gwcode)) %>%
  setNames(c("gwcode", "year", paste0("vdem_", names(.)[3:ncol(.)])))

states <- left_join(states, vdem)
```

## Infant mortality

```{r}
wdi <- read_csv("input/wdi-infmort.csv",
                 col_types = cols(
                   .default = col_double(),
                   year = col_integer(),
                   infmort_imputed = col_logical()
                 )) %>%
  mutate(infmort_imputed = as.integer(infmort_imputed)) %>%
  setNames(c("gwcode", "year", paste0("wdi_", names(.)[3:ncol(.)])))

states <- left_join(states, vdem)
```

## Summarize and write output

```{r final-missplot}
plotmiss(states)
```

Write all incomplete cases to a CSV so changes introduced by something in one of the input datasets is easier to notice:

```{r}
format_years <- function(x) {
  if (length(x) > 1) {
    return(paste(range(x), collapse = " - "))
  }
  as.character(x)
}

incomplete_cases_by_var <- states %>%
  gather(var, value, -gwcode, -year) %>%
  mutate(dv_var = str_detect(var, "lead[0-9]{1}")) %>%
  filter( 
    # for non-DV vars, take any missing values
    (!dv_var & is.na(value)) | 
      # for DV vars, last 2 years are missing by design
      (dv_var & is.na(value) & year < max(states$year) - 1)) %>%
  group_by(gwcode, year, var) %>%
  summarize() 

# the by_var version above is big and inefficient; change it so that we 
# collapse over year ranges, and so that instead of tracking each specific 
# variable, we only identify the data source responsible
incomplete_cases <- incomplete_cases_by_var %>%
  # collapse vars to data source, using the prefix
  mutate(var = str_replace(var, "^([a-z]+)\\_[a-zA-Z0-9\\_]+$", "\\1")) %>%
  group_by(gwcode, year, var) %>%
  summarize() %>%
  # Before we can collapse over year ranges, we need to get the dataset 
  # so that there are no repeated years for a country, otherwise we can't 
  # ID correct [country, missing sources] sets. 
  # Thus, collapse the missing data source info
  group_by(gwcode, year) %>%
  summarize(data_sources = paste0(sort(unique(var)), collapse = "; ")) %>%
  # Collapse year ranges; first we need to id successive year spells so that
  # we don't collapse over gaps, e.g. when a country lost independence during 
  # WW2
  group_by(gwcode) %>%
  arrange(year) %>%
  mutate(spell_id = id_date_sequence(year)) %>%
  # Now collapse over years
  group_by(gwcode, spell_id, data_sources) %>%
  summarize(years = format_years(year),
            N = n()) %>%
  # re-order so it is easier to read
  arrange(gwcode, spell_id) %>%
  select(gwcode, spell_id, years, N, data_sources)
  
write_csv(incomplete_cases, "output/incomplete-cases.csv")
```

### Variables in data

```{r variables}
var_summary <- states %>%
  pivot_longer(everything(), names_to = "variable") %>%
  group_by(variable) %>%
  summarize(missing = sum(is.na(value)),
            sd = sd(value, na.rm = TRUE),
            integer = isTRUE(all.equal(value, as.integer(value))),
            unique_val_ratio = length(unique(value))/length(value))

write_csv(var_summary, "output/variables.csv")

knitr::kable(var_summary, digits = 2)
```


### Missing values by column

```{r summarize-missing-value}
sapply(states, function(x) sum(is.na(x))) %>%
  as.list() %>%
  tibble::enframe(name = "Variable", value = "Missing") %>%
  unnest(Missing) %>%
  filter(Missing > 0) %>%
  knitr::kable()
```

### Track overall cases and missing cases

```{r}
# The DV values for the last year or two will be missing, but those are not
# cases we want to drop as they still serve as prediction targets. So run 
# complete.cases on the data without DVs so we can get an index for rows to 
# drop
no_dv <- states %>%
  select(-ends_with(c("lead1", "lead2")))
drop_idx <- !complete.cases(no_dv)

# Write out high-level summary stats
tbl <- list(
  N_before_drop = nrow(states),
  N_after_drop  = sum(drop_idx==FALSE),
  Years = format_years(states[!drop_idx, ][["year"]]),
  Features = as.integer(ncol(no_dv) - 2),
  Positive_attempt_lead1 = as.integer(
    sum(states[!drop_idx, ][["pt_attempt_lead1"]], na.rm = TRUE)
  ),
  Positive_coup_lead1 = as.integer(
    sum(states[!drop_idx, ][["pt_coup_lead1"]], na.rm = TRUE)
  ),
  Positive_failed_lead1 = as.integer(
    sum(states[!drop_idx, ][["pt_failed_lead1"]], na.rm = TRUE)
  ),
  N_in_forecast_sets = nrow(states[!drop_idx & states$year %in% 2010:2019, ])
)

tbl %>% 
  enframe("Measure", "Value") %>% 
  mutate(Value = as.character(Value)) %>% 
  knitr::kable()

tbl %>%
  yaml::as.yaml() %>%
  writeLines("output/data-summary.yml")
```

### Track forecast sets

```{r}
years         <- 2010:2019
forecast_data <- states[!drop_idx & states$year %in% years, c("gwcode", "year")]

gw_full <- state_panel(min(years), max(years), partial = "any")

# Which countries are covered in each forecast set
covered <- forecast_data %>%
  # ID consecutive year sequences
  group_by(gwcode) %>%
  arrange(gwcode, year) %>%
  mutate(spell = id_date_sequence(year)) %>%
  # collapse over consecutive years
  group_by(gwcode, spell) %>%
  summarize(years = format_years(year)) %>%
  # add country names
  mutate(country = country_names(gwcode, shorten = TRUE)) %>%
  select(gwcode, country, years)
covered %>% 
  knitr::kable(caption = "Countries covered by the test and live forecasts")
write_csv(covered, "output/forecast-covered-countries.csv")

# Which G&W countries are *not* covered by the forecasts
not_covered <- anti_join(gw_full, forecast_data, by = c("gwcode", "year")) %>%
  # ID consecutive year sequences
  group_by(gwcode) %>%
  arrange(gwcode, year) %>%
  mutate(spell = id_date_sequence(year)) %>%
  # collapse over consecutive years
  group_by(gwcode, spell) %>%
  summarize(years = format_years(year)) %>%
  # add country names
  mutate(country = country_names(gwcode, shorten = TRUE)) %>%
  select(gwcode, country, years)
not_covered %>%
  knitr::kable(caption = "Countries *not* covered by the test and live forecasts")
write_csv(not_covered, "output/forecast-not-covered-countries.csv")
```


## Save

```{r}
states_clean <- states[!drop_idx, ]
write_rds(states_clean, "output/states.rds")
```


