---
title: "Ethnic Power Relations (EPR)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Last updated on `r as.character(format(Sys.Date(), "%d %B %Y"))`*

Source: https://icr.ethz.ch/data/epr/core/

License: not clear

Citation: Vogt, Manuel, Nils-Christian Bormann, Seraina Rüegger, Lars-Erik Cederman, Philipp Hunziker, and Luc Girardin. 2015. "Integrating Data on Ethnicity, Geography, and Conflict: The Ethnic Power Relations Data Set Family." Journal of Conflict Resolution 59(7): 1327–42.

To update:

1. Download and get EPR data from the URL above. 

2. Run the `clean-data.Rmd` script (interactively!) and adjust as needed base on new facts on the ground (search for "UPDATE").

3. The main output is "output/epr.csv".

Changes to the data:

- the data are lagged by 2 years in order to have values for 2019
- the resulting initial 2 years of missing values are carry-back imputed, i.e. taken to be the first observed value for a country; otherwise there are no imputations
- some small states are not covered by the data
- several variable transformations are added: binary indicators for whether there have been shifts in the excluded or inpower groups; and year to year changes in the excluded/inpower variables

## Usage

```{r, message=FALSE}
library(readr)
library(dplyr)

epr <- read_csv("output/epr.csv")

glimpse(epr)

range(epr$year)
```

