---
title: "WDI Infant mortality"
output: 
  github_document:
    toc: yes
---

*Last updated on `r as.character(format(Sys.Date(), "%d %B %Y"))`*

Note that places that require attention during data updates are marked with *UPDATE:*

```{r setup-local, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functions / packages

```{r}

library(WDI)
library(tidyverse)
library(states)
library(lubridate)
library(stringr)

# Convert a WDI data frame to G&W statelist
# This will add a gwcode column. It also adds a row for Yugo/Serbia 2006,
# which is why this cannot be a function that creates a G&W column in the 
# existing data frame.
# @param x a data frame genereated by WDI::WDI()
# @returns a data frame like x, but with a gwcode column and extra row
wdi_to_gw <- function(x, iso2c = "iso2c", country = "country", year = "year") {
  
  # In case the ID columns don't match the WDI default, create them here. 
  # dplyr is easier to use when we can refer to these columns directly
  x$iso2c   <- x[[iso2c]]
  x$country <- x[[country]]
  x$year    <- x[[year]]
  
  # remove non-state entities, i.e. regions and aggregates
  notstate <- c("1A", "1W", "4E", "7E", "8S", "B8", "F1", "S1", "S2", "S3", "S4",
                "T2", "T3", "T4", "T5", "T6", "T7", "V1", "V2", "V3", "V4", "Z4",
                "Z7", "EU", "OE", "XC", "XD", "XE", "XF", "XG", "XH", "XI", "XJ",
                "XL", "XM", "XN", "XO", "XP", "XQ", "XT", "XU", "XY", "ZF", "ZG",
                "ZJ", "ZQ", "ZT")
  x <- x %>%
    dplyr::filter(!iso2c %in% notstate) 
  
  # first pass G&W country code coding
  x$gwcode <- suppressWarnings(countrycode::countrycode(x[["iso2c"]], "iso2c", "gwn"))

  # this misses some countries; first the fixed, non-year-varying, cases
  x <- x %>%
    mutate(
      gwcode = case_when(
        iso2c=="AD" ~ 232L,
        iso2c=="XK" ~ 347L,
        country=="Namibia" ~ 565L,
        iso2c=="VN" ~ 816L,  # countrycode uses 817, South Vietnam for this
        iso2c=="YE" ~ 678L,  # Yemen
        TRUE ~ gwcode
      )
    )

  # Fix Serbia/Yugoslavia
  # Right now all coded as 340, but in G&W 345 (Yugo) ends in 2006 and 
  # 340 (Serbia) starts
  serbia2006 <- x[x$gwcode==340 & x$year==2006 & !is.na(x$gwcode), ]
  yugo_idx <- x$gwcode==340 & x$year < 2007 & !is.na(x$gwcode)
  x$gwcode[yugo_idx]  <- 345
  x$iso2c[yugo_idx]   <- "YU"
  x$country[yugo_idx] <- "Yugoslavia/Serbia & Montenegro"
  x <- bind_rows(x, serbia2006) %>%
    arrange(gwcode, iso2c, country, year)
  
  x
}


# I'm going to use this function to impute missing values; see below for more details and the choice of a log-linear model
impute_ts_loglinear <- function(x) {
  # This is only meant to work for leading NAs; check to make sure this is 
  # the case. 
  xna <- is.na(x)
  # if all NA's are leading sequence, they are all preceded by NA as well
  stopifnot(!any(xna==TRUE & c(TRUE, head(stats::lag(xna), -1))==FALSE))
  
  if (all(is.na(x))) {
    return(x)
  }
  xx <- seq_along(x)
  mdl <- try(glm(x ~ xx, data = NULL, family = gaussian(link = "log")), silent = TRUE)
  if (inherits(mdl, "try-error")) {
    return(x)
  }
  xhat <- predict(mdl, newdata = list(xx = xx), type = "response")
  
  # Use the first overlapping point to shift xhat so the lines "connect" instead
  # of having a sudden jump
  pt <- min(which(!xna))
  shift <- x[pt] - xhat[pt]
  xhat  <- xhat + shift
  
  # drop in imputed values
  x[xna] <- xhat[xna]
  x
}

impute_ts_sqrt <- function(x) {
  # This is only meant to work for leading NAs; check to make sure this is 
  # the case. 
  xna <- is.na(x)
  # if all NA's are leading sequence, they are all preceded by NA as well
  stopifnot(!any(xna==TRUE & c(TRUE, head(stats::lag(xna), -1))==FALSE))
  
  if (all(is.na(x))) {
    return(x)
  }
  xx <- seq_along(x)
  mdl <- try(glm(sqrt(x) ~ xx, data = NULL, family = gaussian(link = "identity")), silent = TRUE)
  if (inherits(mdl, "try-error")) {
    return(x)
  }
  xhat <- predict(mdl, newdata = list(xx = xx), type = "response")^2
  
  # Use the first overlapping point to shift xhat so the lines "connect" instead
  # of having a sudden jump
  pt <- min(which(!xna))
  shift <- x[pt] - xhat[pt]
  xhat  <- xhat + shift
  
  # drop in imputed values
  x[xna] <- xhat[xna]
  x
}

# check how this works
x <- c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 
       89.3, 85.1, 81.1, 77.2, 73.5, 70.1, 66.9, 63.9, 61.1, 58.6, 56.2, 
       54.1, 52.1, 50.2, 48.5, 47, 45.8, 44.7, 44.1, 44, 44.4, 45.2, 
       45.9, 46.8, 47.8, 48.4, 49, 49.4, 50, 50.9, 50.8, 49.7, 48.1, 
       45.5, 42.9, 40.3, 38.2, 36.6, 35.3, 34.1, 32.9, 31.4, 30.7, 29.6, 
       28.5)
plot(x, type = "l", ylim = c(0, 150))
lines(impute_ts_loglinear(x)[1:16], col = "red")

```

## Get raw data

Download the raw source data. Since this takes a while, this will only raw if a copy of the source data is not present in the input folder. 

*UPDATE: To trigger and update, delete the raw data input file or manually run this chunk.*

```{r download-source}
if (!file.exists("input/infmort.csv")) {
  raw <- WDI(indicator = "infmort", country = "all", start = 1960, 
           end = year(today()), extra = FALSE)
  write_csv(raw, "input/infmort.csv")
}
```

## Clean raw data

```{r}
raw <- read_csv("input/infmort.csv")

# UPDATE: check this is still the case
# In the raw data all values for 2019 are missing; drop that year from the data
last_year <- filter(raw, year==max(year))
stopifnot(all(is.na(last_year[["SP.DYN.IMRT.IN"]])))
raw <- filter(raw, year!=max(year))

# convert to G&W system
wdi <- raw %>%
  rename(infmort = SP.DYN.IMRT.IN) %>%
  wdi_to_gw(.) %>%
  arrange(gwcode, year)

# Some minor countries don't have G&W codes
nogwcode <- wdi %>%
  filter(is.na(gwcode)) %>%
  group_by(iso2c, country) %>%
  count() 
write_csv(nogwcode, "output/missing-gwcode.csv")
knitr::kable(nogwcode)

# Take those out
wdi <- wdi %>%
  dplyr::filter(!is.na(gwcode))
```

### Lag data

*UPDATE: make sure the lagging is still correct*

Lag the data before we check for an if possible impute missing values because the lagging will introduce missingness as well.

```{r}
wdi$year <- wdi$year + 1
range(wdi$year)
```

### Normalize to G&W statelist

```{r}
# Add in missing cases from G&W state list and drop excess country-years not 
# in G&W list (left join)
statelist <- state_panel(1960,  # this cannot be min(year) because we lagged above
                         max(wdi$year), partial = "any")
wdi <- left_join(statelist, wdi, by = c("gwcode", "year")) %>%
  arrange(gwcode, year)
```

*TODO: this join will take out some excess years for situations where WDI is missing data for a country like Czechoslovakia that later split into several different countries, and for which it does have data. In principle it might be possible to reconstruct mortality estimates for Czechoslovakia, the USSR, Yugoslavia, Sudan, etc. but this is a lot of work, so ¯\_(ツ)_/¯*

## Handle missing values

The rest of this script deals with identifying and potentially imputing missing values. 

```{r}
# Are any years completely missing?
missing_year <- wdi %>%
  group_by(year) %>%
  summarize(n = n(), missing = sum(infmort)) %>%
  filter(n==missing) 
write_csv(missing_year, "output/missing-all-year.csv")
missing_year %>%
  knitr::kable()

# This should not be the case since we lagged the data
stopifnot(nrow(missing_year)==0)

# Are any countries completely missing?
missing_country <- wdi %>%
  group_by(gwcode) %>%
  summarize(n = n(), 
            missing = sum(is.na(infmort))) %>%
  mutate(country = country_names(gwcode, shorten = TRUE)) %>%
  filter(n==missing) %>%
  select(gwcode, country, n)
write_csv(missing_country, "output/missing-all-country.csv")
missing_country %>%
  knitr::kable()

# Take out countries missing all values
wdi <- wdi %>% 
  dplyr::filter(!gwcode %in% missing_country[["gwcode"]])
```

### Carry-back impute lag-induced missing data

Because we lagged the data, countries will have missing values in 1960 or their first year of independence, if it was after 1960. Use the 1960 or independence year value to impute, i.e. carry back impute thoses cases.

```{r}
# Countries that gained indy in 1960 or later
data(gwstates)
indy <- gwstates %>% 
  # some states are present more than 1 time if they had interrupted indy; only
  # use last period
  arrange(gwcode) %>%
  group_by(gwcode) %>% 
  slice(n()) %>%
  mutate(syear = lubridate::year(start)) %>%
  select(gwcode, syear)
stopifnot(nrow(indy)==length(unique(indy$gwcode)))

wdi <- wdi %>%
  left_join(indy, by = "gwcode") %>%
  group_by(gwcode) %>%
  arrange(gwcode, year) %>%
  mutate(infmort2 = case_when(
    (is.na(infmort) & year>=syear) ~ lead(infmort, n = 1)[1],
    TRUE ~ infmort
  ))

# I add the imputed values as second column to allow comparison, if one wants
# to do that at this point. 
sum(is.na(wdi$infmort))
sum(is.na(wdi$infmort2))

wdi <- wdi %>%
  mutate(infmort = infmort2, infmort2 = NULL)
```

### Check remaining missing values

```{r}
missing <- wdi %>%
  # Track the original N for a country before focusing only on missing cases
  group_by(gwcode) %>%
  dplyr::mutate(N = n()) %>%
  filter(is.na(infmort)) %>%
  # Right now this is country-year; turn this into a more readable form by 
  # collapsing accross consecutive year spells
  # First, code consecutive year spells
  group_by(gwcode) %>%
  arrange(year) %>%
  dplyr::mutate(year = as.integer(year),
                id = id_date_sequence(year)) %>%
  # Collapse over years
  dplyr::group_by(gwcode, id) %>%
  dplyr::summarize(N = unique(N), 
                   N_miss = n(),
                   Frac_miss = N_miss/N,
                   years = paste0(range(year), collapse = " - ")) %>%
  select(-id) %>%
  arrange(desc(Frac_miss), gwcode)

missing %>% 
  knitr::kable()

# add an indicator if series is incomplete 
wdi <- wdi %>%
  mutate(has_missing = gwcode %in% missing$gwcode)
```

These are all missing values at the front of the series. 

### Find imputation model

The series overall look like the might be reasonably linear under some scale transformation like log or square root. 

```{r}
ggplot(wdi, aes(x = year, y = infmort, group = gwcode, 
                color = has_missing)) +
  geom_line(alpha = 0.5) +
  theme_light()
```

Compare log-linear and square root linear models.

```{r}
fit <- wdi %>%
  group_by(gwcode, has_missing) %>%
  nest() %>%
  mutate(
    mdl_log = map(data, ~lm(log(infmort) ~ year, data = .)),
    mdl_sqrt = map(data, ~lm(sqrt(infmort) ~ year, data = .)),
    mdl_mix  = map(data, ~lm((infmort)^(0.42) ~ year, data = .))
  ) %>%
  gather(model, fit, starts_with("mdl")) %>%
  mutate(r2 = map_dbl(fit, ~summary(.)$r.squared))

# Bin the fit by R^2 and model
table(fit$model, cut(fit$r2, c(0, .4, .5, .6, .7, .8, .9, 1)))


fit %>% 
  group_by(model, has_missing) %>%
  summarize(countries = n(),
            mean_r2 = round(mean(r2), 2),
            median_r2 = round(median(r2), 2)) %>%
  arrange(has_missing, mean_r2)
```

If a model is not performing well on a series where wer are not looking to impute, who cares. Look at low R2 models for series we are looking to impute.

```{r}
check_gwcodes <- fit %>% 
  filter(model!="mdl_lm", r2 < 0.9, has_missing) %>% 
  pull(gwcode) %>% 
  unique()
bad_fit <- fit %>%
  filter(gwcode %in% check_gwcodes, model!="mdl_lm") 
nn <- nrow(bad_fit)/3
wdi %>%
  filter(gwcode %in% check_gwcodes) %>%
  mutate(syear = ifelse(syear<1960, 1960, syear)) %>%
  ggplot(.) +
  facet_wrap(~ gwcode) +
  geom_line(aes(x = year, y = infmort, group = gwcode)) +
  # mark how far back we need to impute
  geom_vline(aes(xintercept = syear)) +
  geom_text(
    data = bad_fit, 
    x = 1980, 
    y = c(rep(50, nn), rep(100, nn), rep(150, nn)), 
    aes(label = paste0(
      c(rep("Log-linear: ", nn), rep("Quadratic: ", nn), 
      rep("Mixed: ", nn)),
      round(r2, 2))
    )
  ) +
  stat_smooth(aes(x = year, y = infmort, color = "red"),
              method = "glm", 
              formula = y ~ x, 
              method.args = list(family = gaussian(link = "log"))) +
  stat_smooth(aes(x = year, y = infmort, color = "blue"),
              method = "lm",
              formula = y ~ poly(x, 2)) +
  scale_color_manual("Model", values = c("red" = "red", "blue" = "blue"), 
                     labels = c("blue" = "Quadratic", "red" = "Log-linear")) +
  theme_light()
```

The log-lienar and square root models both perform about equally well. What do the imputed values look like?


```{r}
wdi <- wdi %>%
  group_by(gwcode) %>%
  arrange(gwcode, year) %>%
  mutate(infmort_imputed = is.na(infmort),
         infmort_log     = impute_ts_loglinear(infmort),
         infmort_sqrt    = impute_ts_sqrt(infmort))

# Visualize results
highlight <- wdi %>% 
  filter(has_missing) %>%
  select(gwcode, year, infmort_imputed, infmort_log, infmort_sqrt) %>%
  pivot_longer(infmort_log:infmort_sqrt) %>%
  mutate(color = case_when(
    infmort_imputed==FALSE ~ "Observed",
    infmort_imputed==TRUE & name=="infmort_log" ~ "Log",
    infmort_imputed==TRUE & name=="infmort_sqrt" ~ "Sqrt",
    TRUE ~ "what?"
  ))
ggplot(wdi, aes(x = year)) +
  geom_line(aes(y = infmort, group = gwcode), alpha = 0.2) +
  geom_line(data = highlight,
            aes(y = value, group = interaction(gwcode, name), color = color)) +
  scale_color_manual(values = c("Observed" = "gray20", "Log" = "red", "Sqrt" = "blue")) +
  theme_light()
```

The square root model's imputed values are less aggressive in their extrapolation so I will pick that.

```{r}
wdi <- wdi %>%
  mutate(infmort = infmort_sqrt,
         infmort_log = NULL)
```

Now all the countries we have not dropped should have full series / no missing values. 
```{r check-no-missing-left}
tbl <- wdi %>%
  group_by(gwcode) %>%
  summarize(n_miss = sum(is.na(infmort))) %>%
  ungroup() %>%
  filter(n_miss > 0)
if (nrow(tbl) > 0) stop("Still some missing values remaining", call. = FALSE)
```


## Add year-normalized version

```{r yearadj-version}
wdi <- wdi %>%
  group_by(year) %>%
  mutate(infmort_yearadj = (infmort - mean(infmort))/sd(infmort)) %>%
  ungroup()
```

## Done, save

```{r}
wdi <- wdi %>% 
  ungroup() %>%
  select(gwcode, year, infmort, infmort_yearadj, infmort_imputed) 
write_csv(wdi, path = "output/wdi-infmort.csv")
```

