---
title: "V-Dem"
output: 
  github_document:
    toc: yes
---

*Last updated on `r as.character(format(Sys.Date(), "%d %B %Y"))`*

This script:

- normalizes the V-Dem data to the G&W statelist
- subsets the V-Dem core indices starting with "v2x_" only
- imputes almost all missing values for countries with partial missing data using a combination of linear imputation for gaps and carry-back for leading sequences of NAs; countries that are entirely missing are not imputed
- adds year to year diff ("_d1") and 5-year moving SD ("_sd") variable transformations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages / functions

```{r}
library(readr)
library(states)
library(dplyr)
library(imputeTS)
library(tidyr)
library(ggplot2)

# There are two patterns of missing values in the data here. One are short gaps,
# for which it is more reasonable to use linear imputation. The other are long
# series of missing values at the beginning of a series. Carry-back the first
# observed value for those. 
# 
# I'm going to use imputesTS::na_interpolation(method = "linear") to impute
# values. This uses stats::approx(rule = 2) under the hood, which almost does 
# exactly what I need: 
#   - for NA values that form gaps, i.e. are surrounded by non-missing values
#     on both sides, linear interpolation is used
#   - for NA values at the head or tail of a series, i.e. which only have a 
#     non-missing value on one side, use that value.
# I actually don't want to extrapolate missing values at the tail, e.g. if I 
# have data through 2018 and need a value for 2019. The imputer() wrapper
# below just makes sure I don't do that for ending NA values. 

# The first two functions below implement a version of rle that treats 
# consecutive NA values as equal. This is used to ID whether a sequence of 
# missing values is a gap or whether it is at the head or tail of a series. 

# Version of `==` for which NA==NA is TRUE and NA==1 is FALSE 
na_equal <- function(x, y) {
  stopifnot(is.vector(x),
            is.vector(y),
            length(x)==length(y))
  z <- logical(length(x))
  for (i in seq_along(x)) {
    if (is.na(x[i]) & is.na(y[i])) z[i] <- TRUE; next
    z[i] <- x[i]==y[i]
  }
  z
}

# Version of rle in which missing values are regarded as equal
rle_na <- function (x) {
  if (!is.vector(x) && !is.list(x)) 
    stop("'x' must be a vector of an atomic type")
  n <- length(x)
  if (n == 0L) 
    return(structure(list(lengths = integer(), values = x), 
                     class = "rle"))
  # the only change
  #y <- x[-1L] != x[-n]
  y <- !na_equal(x[-1L], x[-n])
  i <- c(which(y | is.na(y)), n)
  structure(list(lengths = diff(c(0L, i)), values = x[i]), 
            class = "rle")
}

imputer <- function(x) {
  
  # escape na_interpolation error
  if (sum(!is.na(x)) < 3) {
    return(x)
  }
  
  xrle <- rle_na(x)
  xhat <- imputeTS::na_interpolation(x, option = "linear")
  if (is.na(tail(xrle$values, 1))) {
    tt <- tail(xrle$lengths, 1)
    n  <- length(xhat)
    xhat[(n - tt + 1):n] <- NA
  }
  xhat
}


# Check to make sure it's working; the final 2 NA values should still be NA
x <- c(rep(NA, 10), 3, 3, NA, NA, 6, 6, 6, rep(NA, 2))
xhat <- imputer(x)
stopifnot(all(is.na(xhat[18:19])))
plot(x, type = "l")
points(xhat, col = "red")
```

## Clean raw data

```{r clean-raw-data}
raw = read_csv("input/v10/V-Dem-CY-Core-v10.csv",
               col_types = cols(
                 gapend1 = col_integer(),
                 gapend2 = col_integer(),
                 gapend3 = col_integer(),
                 gapstart1 = col_integer(),
                 gapstart2 = col_integer(),
                 gapstart3 = col_integer()
               ))

v2x = raw %>%
  # Most v-dem incies starts with v2x, but some have the form v2x[a-z]{2}
  # Thus uses machtes() rather than starts_with()
  select(country_name, COWcode, year, matches("^v2x[a-z]{0,2}\\_", perl = TRUE)) %>%
  # We don't need to credible intervals, sd, etc.
  select(-ends_with("codehigh"), -ends_with("codelow"), -ends_with("sd"),
         -ends_with("osp")) %>%
  # take out the discrete regime variable, it will get treated as integer
  # and i don't want to deal with dummying it out
  select(-starts_with("v2x_regime")) %>%
  # Variables with lots of missing values
  # UPDATE: run this first with them and see below whether they still have 
  # missing value issues
  # academic freedom index 
  select(-v2xca_academ) %>%
  # legislative control of the executive
  select(-v2xlg_legcon) %>%
  # party institutionalization
  select(-v2xps_party) %>%
  # local government index
  select(-v2xel_locelec) %>%
  # political exclusion
  select(-contains("v2xpe")) %>%  # completely missing for some countries
  # direct democracy
  select(-contains("v2xdd")) %>%  # completely missing for some countries
  # 2020-04-17: imputing all these indices is going to be a nightmare;
  # for now revert back to only v2x_ versions
  select(country_name, COWcode, year, starts_with("v2x_")) %>%
  # G&W doesn't start until 1816, so take out earlier
  filter(year > 1815)
```

### Normalize to G&W statelist

I'm only concerned about getting data for 1960 on. Further below I'm adding a year to year diff, so make it 1959 to avoid missing first year values for those transformations.

```{r normalize-to-gw}
# Serbia/Yugoslavia change in 2006
# Change 345 to 340 from 2006 on, below. But then we also need to add a row for 3
# 45 in 2006, with same values. Extract this now.
yugo2006 <- v2x %>% filter(year==2006 & COWcode==345) %>%
  rename(gwcode = COWcode)

# Change from COW codes to G&W codes
v2xgw = v2x %>%
  rename(gwcode = COWcode) %>%
  mutate(gwcode = case_when(
    gwcode == 255 ~ 260,
    gwcode == 679 ~ 678,
    gwcode == 345 & year >= 2006 ~ 340,
    TRUE ~ gwcode))

# add Yugo 2006 back in
v2xgw <- bind_rows(v2xgw, yugo2006) %>%
  arrange(gwcode, country_name, year)

# Keep only post-1959 data
v2xgw <- filter(v2xgw, year >= 1959)

# Some non-state entities are missing GW codes, drop them
v2xgw %>% 
  filter(is.na(gwcode)) %>%
  group_by(country_name) %>%
  summarize(years = paste0(range(year), collapse = " - "))
v2xgw <- filter(v2xgw, !is.na(gwcode))

# Check case-list inconsistencies; this can be helpful for ID'ing country code
# mismatches
# UPDATE: check both instances of "check" below manually
statelist <- state_panel(1959, max(v2xgw$year), partial = "any")
sets <- states::compare(v2xgw, statelist)

data(gwstates)
gwstates <- gwstates %>%
  select(gwcode, country_name, start, end) %>%
  # some states have more than one spell; only check last
  group_by(gwcode) %>%
  slice(n())

# Cases in v2xgw but not in statelist; these could be miscoded GW codes
check <- sets %>% 
  filter(case_in_df1, !case_in_df2) %>%
  left_join(gwstates, by = "gwcode") %>%
  group_by(gwcode) %>%
  summarize(years = paste0(range(year), collapse = " - "),
            start = unique(start), end = unique(end))

# Cases in statelist but not in v2xgw
check <- sets %>% 
  filter(case_in_df2, !case_in_df1) %>%
  left_join(gwstates, by = "gwcode") %>%
  group_by(gwcode) %>%
  summarize(years = paste0(range(year), collapse = " - "),
            country_name = unique(country_name),
            start = unique(start), end = unique(end))

# Add in missing cases from G&W state list and drop excess country-years not 
# in G&W list (left join)
v2xgw <- left_join(statelist, v2xgw, by = c("gwcode", "year")) %>%
  arrange(gwcode, year)
```

## Handle missing values

```{r handle-missing-values}
# Switch names again to make it shorter; doing it here allows me to re-run the
# chunk above if something goes wrong here during interactive checking
v2x <- v2xgw

# Are any countries completely missing?
missing_country <- v2x %>%
  group_by(gwcode) %>%
  # track original number of years for gwcode
  mutate(n = n()) %>%
  # filter out so we only keep rows with missing values for ALL non-ID cols
  filter_at(vars(starts_with("v2x")), all_vars(is.na(.))) %>%
  group_by(gwcode, n) %>%
  summarize(missing = n()) %>%
  mutate(country = country_names(gwcode, shorten = TRUE)) %>%
  filter(n==missing) %>%
  select(gwcode, country, n)

write_csv(missing_country, "output/missing-all-country.csv")
missing_country %>%
  knitr::kable()

# Take out countries missing all values
v2x <- v2x %>% 
  dplyr::filter(!gwcode %in% missing_country[["gwcode"]])

# Now check whether any particular variable is missing completely for a year
# I'm not ok imputing in such an instance, so in that case the variable should
# be taken out.
check <- v2x %>% 
  group_by(gwcode) %>% 
  mutate(n = n(), year = NULL) %>% 
  group_by(gwcode, n, country_name) %>% 
  # for each column we will now have # of missing vals
  summarize_all(~sum(is.na(.))) %>%
  # only check ones with at least 1 missing value
  filter_at(vars(starts_with("v2x")), any_vars(. > 0)) %>%
  # select only vars that have missing vals
  pivot_longer(-c(gwcode, n, country_name)) %>%
  group_by(name) %>%
  mutate(any_na = any(value > 0)) %>%
  filter(any_na) %>%
  select(-any_na) %>%
  pivot_wider()

# How many missing values are there for each variable?
sapply(v2x, function(x) sum(is.na(x))) %>%
  tibble::enframe(name = "variable", value = "missing") %>%
  arrange(desc(missing)) %>%
  filter(missing > 0) %>%
  knitr::kable()

#
#   Start imputing missing series, one by one
#   __________________________________________
# 
#   UPDATE: all of these will need manual rechecking during data updates
#
#   First, retain a copy of orig v2x so we can later mark imputed values
#

v2x_orig <- v2x %>%
  select(-country_name) %>%
  setNames(c("gwcode", "year", paste0("orig_", names(.)[3:ncol(.)])))

# Bahrain is missing the early parts for several series
filter(v2x, gwcode==692) %>%
  pivot_longer(-c(gwcode, year, country_name)) %>%
  # only keep variables with missing vals
  group_by(name) %>%
  mutate(any_na = any(is.na(value))) %>%
  ungroup() %>%
  filter(any_na) %>%
  ggplot(aes(x = year, y = value, group = name)) +
  geom_line(data = v2x %>% filter(gwcode==692) %>% 
              select(-gwcode, -country_name) %>% pivot_longer(-year),
  alpha = 0.2) +
  geom_line(aes(color = name))

stopifnot(all(is.na(
  filter(v2x, gwcode==692, year < 2002) %>%
    select(v2x_libdem, v2x_liberal, v2x_jucon, v2x_corr) %>%
    unlist()
)))

# carry back first observed value
v2x[v2x$gwcode==692, ] <- v2x[v2x$gwcode==692, ] %>% 
  tidyr::fill(-c(gwcode, year, country_name), .direction = "up")

# Cameroon is missing first year values; take 2nd year values
stopifnot(all(is.na(v2x %>% filter(gwcode==471 & year==1960) %>% select(-gwcode, -year))))
drop_in <- v2x %>% filter(gwcode==471 & year==1961) %>% mutate(year = 1960)
v2x[v2x$gwcode==471 & v2x$year==1960, ] <- drop_in

# Mozambique; missing several indices during civil war years; carry back first 
# available value
filter(v2x, gwcode==541) %>%
  pivot_longer(-c(gwcode, year, country_name)) %>%
  # only keep variables with missing vals
  group_by(name) %>%
  mutate(any_na = any(is.na(value))) %>%
  ungroup() %>%
  filter(any_na) %>%
  ggplot(aes(x = year, y = value, group = name)) +
  geom_line(data = v2x %>% filter(gwcode==692) %>% 
              select(-gwcode, -country_name) %>% pivot_longer(-year),
  alpha = 0.2) +
  geom_line(aes(color = name))

stopifnot(all(is.na(
  filter(v2x, gwcode==541, year==1975) %>% select(v2x_polyarchy, v2x_libdem, v2x_partipdem,
                                      v2x_delibdem, v2x_api) %>%
    unlist()
)))

v2x[v2x$gwcode==541, ] <- v2x[v2x$gwcode==541, ] %>% 
  tidyr::fill(-c(gwcode, year, country_name), .direction = "up")

# CAR; has a couple of gaps, so linear impute
filter(v2x, gwcode==482) %>%
  pivot_longer(-c(gwcode, year, country_name)) %>%
  # only keep variables with missing vals
  group_by(name) %>%
  mutate(any_na = any(is.na(value))) %>%
  ungroup() %>%
  filter(any_na) %>%
  ggplot(aes(x = year, y = value, group = name)) +
  geom_line(data = v2x %>% filter(gwcode==692) %>% 
              select(-gwcode, -country_name) %>% pivot_longer(-year),
  alpha = 0.2) +
  geom_line(aes(color = name))

stopifnot(all(is.na(
  filter(v2x, gwcode==482, year %in% c(1964, 1965)) %>%
    select(v2x_libdem, v2x_liberal, v2x_gender, v2x_genpp) %>%
    unlist()
)))

filter(v2x, gwcode==482) %>% 
  select(year, v2x_libdem) %>% 
  plot(main = "CAR", type = "b")
v2x$v2x_libdem[v2x$gwcode==482] <- na_interpolation(v2x$v2x_libdem[v2x$gwcode==482], "linear")
v2x$v2x_liberal[v2x$gwcode==482] <- na_interpolation(v2x$v2x_liberal[v2x$gwcode==482], "linear")
v2x$v2x_gender[v2x$gwcode==482] <- na_interpolation(v2x$v2x_gender[v2x$gwcode==482], "linear")
v2x$v2x_genpp[v2x$gwcode==482] <- na_interpolation(v2x$v2x_genpp[v2x$gwcode==482], "linear")

# Saudi Arabia is missing 2 gender vars for early part of series
filter(v2x, gwcode==670) %>%
  pivot_longer(-c(gwcode, year, country_name)) %>%
  # only keep variables with missing vals
  group_by(name) %>%
  mutate(any_na = any(is.na(value))) %>%
  ungroup() %>%
  filter(any_na) %>%
  ggplot(aes(x = year, y = value, group = name)) +
  geom_line(data = v2x %>% filter(gwcode==670) %>% 
              select(-gwcode, -country_name) %>% pivot_longer(-year),
  alpha = 0.2) +
  geom_line(aes(color = name))

stopifnot(all(is.na(
  filter(v2x, gwcode==670, year < 2000) %>% select(v2x_gender, v2x_genpp) %>% unlist()
)))

v2x[v2x$gwcode==670, ] <- v2x[v2x$gwcode==670, ] %>% 
  tidyr::fill(-c(gwcode, year, country_name), .direction = "up")

# Qatar is also missing a lot for 2 gender vars
filter(v2x, gwcode==694) %>%
  pivot_longer(-c(gwcode, year, country_name)) %>%
  # only keep variables with missing vals
  group_by(name) %>%
  mutate(any_na = any(is.na(value))) %>%
  ungroup() %>%
  filter(any_na) %>%
  ggplot(aes(x = year, y = value, group = name)) +
  geom_line(data = v2x %>% filter(gwcode==694) %>% 
              select(-gwcode, -country_name) %>% pivot_longer(-year),
  alpha = 0.2) +
  geom_line(aes(color = name))

stopifnot(all(is.na(
  filter(v2x, gwcode==694, year < 2002) %>%
    select(v2x_gender, v2x_genpp) %>%
    unlist()
)))

# carry back first observed value
v2x[v2x$gwcode==694, ] <- v2x[v2x$gwcode==694, ] %>% 
  tidyr::fill(-c(gwcode, year, country_name), .direction = "up")

# Singapore is missing 2019 v2x_feduni (division of powers index) value
# this will not get imputed below since it is at the tail of a series; all other
# values are 0, so safe to use this again
v2x %>% 
  filter(gwcode==830) %>%
  pull(v2x_feduni) %>% 
  table()

stopifnot(is.na(v2x$v2x_feduni[v2x$gwcode==830 & v2x$year==2019]))
v2x$v2x_feduni[v2x$gwcode==830 & v2x$year==2019] <- 0


# 
#   The remaining gaps are all N < 20; apply the same imputation logic to 
#   all of them
#   _______________________________
#
#   UPDATE: make sure in check below that there are no cells with > 20
#

check <- v2x %>% 
  group_by(gwcode) %>% 
  mutate(n = n(), year = NULL) %>% 
  group_by(gwcode, n, country_name) %>% 
  # for each column we will now have # of missing vals
  summarize_all(~sum(is.na(.))) %>%
  # only check ones with at least 1 missing value
  filter_at(vars(starts_with("v2x")), any_vars(. > 0)) %>%
  # select only vars that have missing vals
  pivot_longer(-c(gwcode, n, country_name)) %>%
  group_by(name) %>%
  mutate(any_na = any(value > 0)) %>%
  filter(any_na) %>%
  select(-any_na) %>%
  pivot_wider()
check

v2x <- v2x %>%
  group_by(gwcode) %>%
  arrange(gwcode, year) %>%
  mutate_at(vars(-c(gwcode, year, country_name)), imputer)

# How many missing values are there for each variable?
sapply(v2x, function(x) sum(is.na(x))) %>%
  tibble::enframe(name = "variable", value = "missing") %>%
  arrange(desc(missing)) %>%
  filter(missing > 0) %>%
  knitr::kable()

# Mark imputed values 
# This is some helper code below to set up this big blog of text
# cat(paste0(names(v2x), "_imputed = as.integer(!is.na(", names(v2x), ") &\n is.na(orig_", names(v2x), "))"), sep = ",\n")
# 
# There's probably a way to do this with mutate_at but I'm not going to mess 
# around with quosures
v2x <- left_join(v2x, v2x_orig, by = c("gwcode", "year")) %>%
  mutate(
    v2x_polyarchy_imputed = as.integer(!is.na(v2x_polyarchy) &
                                         is.na(orig_v2x_polyarchy)),
    v2x_libdem_imputed = as.integer(!is.na(v2x_libdem) &
                                      is.na(orig_v2x_libdem)),
    v2x_partipdem_imputed = as.integer(!is.na(v2x_partipdem) &
                                         is.na(orig_v2x_partipdem)),
    v2x_delibdem_imputed = as.integer(!is.na(v2x_delibdem) &
                                        is.na(orig_v2x_delibdem)),
    v2x_egaldem_imputed = as.integer(!is.na(v2x_egaldem) &
                                       is.na(orig_v2x_egaldem)),
    v2x_api_imputed = as.integer(!is.na(v2x_api) &
                                   is.na(orig_v2x_api)),
    v2x_mpi_imputed = as.integer(!is.na(v2x_mpi) &
                                   is.na(orig_v2x_mpi)),
    v2x_freexp_altinf_imputed = as.integer(!is.na(v2x_freexp_altinf) &
                                             is.na(orig_v2x_freexp_altinf)),
    v2x_frassoc_thick_imputed = as.integer(!is.na(v2x_frassoc_thick) &
                                             is.na(orig_v2x_frassoc_thick)),
    v2x_suffr_imputed = as.integer(!is.na(v2x_suffr) &
                                     is.na(orig_v2x_suffr)),
    v2x_elecoff_imputed = as.integer(!is.na(v2x_elecoff) &
                                       is.na(orig_v2x_elecoff)),
    v2x_liberal_imputed = as.integer(!is.na(v2x_liberal) &
                                       is.na(orig_v2x_liberal)),
    v2x_jucon_imputed = as.integer(!is.na(v2x_jucon) &
                                     is.na(orig_v2x_jucon)),
    v2x_partip_imputed = as.integer(!is.na(v2x_partip) &
                                      is.na(orig_v2x_partip)),
    v2x_cspart_imputed = as.integer(!is.na(v2x_cspart) &
                                      is.na(orig_v2x_cspart)),
    v2x_egal_imputed = as.integer(!is.na(v2x_egal) &
                                    is.na(orig_v2x_egal)),
    v2x_accountability_imputed = as.integer(!is.na(v2x_accountability) &
                                              is.na(orig_v2x_accountability)),
    v2x_veracc_imputed = as.integer(!is.na(v2x_veracc) &
                                      is.na(orig_v2x_veracc)),
    v2x_diagacc_imputed = as.integer(!is.na(v2x_diagacc) &
                                       is.na(orig_v2x_diagacc)),
    v2x_horacc_imputed = as.integer(!is.na(v2x_horacc) &
                                      is.na(orig_v2x_horacc)),
    v2x_ex_confidence_imputed = as.integer(!is.na(v2x_ex_confidence) &
                                             is.na(orig_v2x_ex_confidence)),
    v2x_ex_direlect_imputed = as.integer(!is.na(v2x_ex_direlect) &
                                           is.na(orig_v2x_ex_direlect)),
    v2x_ex_hereditary_imputed = as.integer(!is.na(v2x_ex_hereditary) &
                                             is.na(orig_v2x_ex_hereditary)),
    v2x_ex_military_imputed = as.integer(!is.na(v2x_ex_military) &
                                           is.na(orig_v2x_ex_military)),
    v2x_ex_party_imputed = as.integer(!is.na(v2x_ex_party) &
                                        is.na(orig_v2x_ex_party)),
    v2x_neopat_imputed = as.integer(!is.na(v2x_neopat) &
                                      is.na(orig_v2x_neopat)),
    v2x_civlib_imputed = as.integer(!is.na(v2x_civlib) &
                                      is.na(orig_v2x_civlib)),
    v2x_clphy_imputed = as.integer(!is.na(v2x_clphy) &
                                     is.na(orig_v2x_clphy)),
    v2x_clpol_imputed = as.integer(!is.na(v2x_clpol) &
                                     is.na(orig_v2x_clpol)),
    v2x_clpriv_imputed = as.integer(!is.na(v2x_clpriv) &
                                      is.na(orig_v2x_clpriv)),
    v2x_corr_imputed = as.integer(!is.na(v2x_corr) &
                                    is.na(orig_v2x_corr)),
    v2x_execorr_imputed = as.integer(!is.na(v2x_execorr) &
                                       is.na(orig_v2x_execorr)),
    v2x_pubcorr_imputed = as.integer(!is.na(v2x_pubcorr) &
                                       is.na(orig_v2x_pubcorr)),
    v2x_gender_imputed = as.integer(!is.na(v2x_gender) &
                                      is.na(orig_v2x_gender)),
    v2x_gencl_imputed = as.integer(!is.na(v2x_gencl) &
                                     is.na(orig_v2x_gencl)),
    v2x_gencs_imputed = as.integer(!is.na(v2x_gencs) &
                                     is.na(orig_v2x_gencs)),
    v2x_genpp_imputed = as.integer(!is.na(v2x_genpp) &
                                     is.na(orig_v2x_genpp)),
    v2x_rule_imputed = as.integer(!is.na(v2x_rule) &
                                    is.na(orig_v2x_rule)),
    v2x_elecreg_imputed = as.integer(!is.na(v2x_elecreg) &
                                       is.na(orig_v2x_elecreg)),
    v2x_EDcomp_thick_imputed = as.integer(!is.na(v2x_EDcomp_thick) &
                                            is.na(orig_v2x_EDcomp_thick)),
    v2x_freexp_imputed = as.integer(!is.na(v2x_freexp) &
                                      is.na(orig_v2x_freexp)),
    v2x_hosabort_imputed = as.integer(!is.na(v2x_hosabort) &
                                        is.na(orig_v2x_hosabort)),
    v2x_hosinter_imputed = as.integer(!is.na(v2x_hosinter) &
                                        is.na(orig_v2x_hosinter)),
    v2x_legabort_imputed = as.integer(!is.na(v2x_legabort) &
                                        is.na(orig_v2x_legabort)),
    v2x_divparctrl_imputed = as.integer(!is.na(v2x_divparctrl) &
                                          is.na(orig_v2x_divparctrl)),
    v2x_feduni_imputed = as.integer(!is.na(v2x_feduni) &
                                      is.na(orig_v2x_feduni))
  ) %>% 
  select(-starts_with("orig_"))
```

## Add variable transformations

```{r variable-transformations}
#
#   Add variable transformations ----
#   ___________________________ 

# first we need to identify country-spells, to make sure gaps don't accidentally
# cross over into what are supposed to be year to year changes
v2x = v2x %>%
  group_by(gwcode) %>%
  arrange(year) %>%
  mutate(spell_id = id_date_sequence(year))

# modified version of rolling SD with 0 values for first value in series 
# instead of NA
rolling_sd <- function(x) {
  z <- zoo::rollapplyr(x, FUN = sd, width = 5, partial = TRUE)
  if (!is.na(x[1])) z[1] <- 0
  z
}

v2x_feats = v2x %>%
  group_by(gwcode, spell_id) %>%
  arrange(gwcode, spell_id, year) %>% 
  mutate_at(vars(-c(gwcode, year, spell_id, country_name, ends_with("imputed"))), 
            list(d1 = ~c(0, diff(.)), 
                 sd = ~rolling_sd(.))
            )



v2x <- v2x_feats
```


## Done, save

```{r}
# Take out 1959, which we had to avoid initial first year missing values, e.g.
# for the diff transformations above. 
v2x <- v2x %>% 
  ungroup() %>%
  filter(year > 1959) %>%
  select(-spell_id)

write_csv(v2x, "output/v-dem.csv")
```

