---
title: "2020 forecasts"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE)

library(leaflet)
library(leaflet.providers)
library(sf)
```

Forecasts for the risk of a successful coup ("Successful Coup"), failed coup attempt ("Failed Attempt") and any coup attempt, regardless of outcome ("Any Coup Attempt"), in 2020. 

The risk values shown are percentages that can theoretically range from 0% to 100%. These are created from the underlying raw probabilities, which can range from 0 to 1, and then rounded to 1 decimal point. 

```{r, include = TRUE}
World <- readRDS("_data/World.rds")

# x is the input probability, 0-1
format_prob <- function(x) {
  perc <- sprintf("%.1f%%", x)
  # Cases that round to 0.0
  perc[x < 0.05 & !is.na(x)] <- "<0.1%"
  perc[is.na(x)] <- "n/a"
  perc
}

label_html <- "
<strong>%s</strong>
  <div class = 'leaflet-popup-body'>
    <p>
      <span>Attempt</span>
      <span></span>
      <span>%s</span>
    </p>
    <p>
      <span>Coup</span>
      <span></span>
      <span>%s</span>
    </p>
    <p>
      <span>Failed</span>
      <span></span>
      <span>%s</span>
    </p>
  </div>
"

labels = sprintf(label_html,
  World$name, format_prob(World$attempt), format_prob(World$coup),
  format_prob(World$failed)
) %>% lapply(htmltools::HTML)

# I previously used colorQuantile for this, but this distorts how high the risk
# really is in some countries. Better to use explicit bins. 
pal_p <- colorBin("viridis", bins = c(0, 0.5, 1, 2, 5, Inf))

leaflet(width = "100%", height = "550px", options = leafletOptions(maxZoom = 5, minZoom = 2)) %>%
  #addProviderTiles("Esri.WorldGrayCanvas", group = "Esri.WorldGrayCanvas") %>%
  addPolygons(data = World, weight = 1, color = "black", smoothFactor = 0.3,
              fillOpacity = 0.8, fillColor = ~pal_p(attempt),
              popup = labels,
              group = "Any Coup Attempt",
              highlight = highlightOptions(weight = 3)
              ) %>%
  # coup
  addPolygons(data = World, weight = 1, color = "black", smoothFactor = 0.3,
              fillOpacity = 0.8, fillColor = ~pal_p(coup),
              popup = labels,
              group = "Successful Coup",
              highlight = highlightOptions(weight = 3)
              ) %>%
  # failed attempt
  addPolygons(data = World, weight = 1, color = "black", smoothFactor = 0.3,
              fillOpacity = 0.8, fillColor = ~pal_p(failed),
              popup = labels,
              group = "Failed Attempt",
              highlight = highlightOptions(weight = 3)
              ) %>% 
  addLegend(pal = pal_p, values = World$coup, opacity = 0.8,
            labFormat = labelFormat(big.mark = "'"),
            title = "Statistical Risk (%)") %>%
  addLayersControl(
    baseGroups = c("Any Coup Attempt", "Successful Coup", "Failed Attempt"),
                   position = "topleft",
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  setView(lng = 0, lat = 20, zoom = 2) 
```

You can download these forecasts [here (CSV)](https://raw.githubusercontent.com/andybega/forecaster2/master/forecasts/coup-2020.csv).

