---
title: "Table of forecasts"
---

The table below lists all countries for which there are forecasts. The columns correspond to different outcomes:

- Attempt: Any coup attempt, regardless of whether it was successful or failed. 
- Coup: A successful coup that results in overthrow of the government.
- Failed: A failed coup attempt that does not result in overthrow of the government.

The risk values shown are percentages that can theoretically can range from 0% to 100%. These were created from raw risk probabilities that range from 0 to 1 and then are rounded to one decimal point. 

```{r setup, include = FALSE}
library(dplyr)
library(DT)
library(states)
library(tidyr)
library(stringr)
```


```{r, echo = FALSE}
fcasts <- readRDS("_data/fcasts.rds") 
fcasts <- fcasts %>%
  # take out the tests forecasts so we only retain the live forecast
  filter(year==max(year)) %>% 
  mutate(Country = country_names(gwcode),
         Country = str_trim(str_remove(Country, " \\([a-zA-Z ]+\\)")),
         Country = str_remove(Country, "/[a-zA-Z]+"),
         Country = case_when(
           Country=="German Federal Republic" ~ "Germany",
           TRUE ~ Country
         )) %>%
  # capitalize first letter in outcome names
  mutate(outcome = str_to_sentence(outcome)) %>%
  pivot_wider(names_from = outcome, values_from = p) %>%
  mutate(
    Attempt_Rank = n() - rank(Attempt) + 1,
    Coup_Rank    = n() - rank(Coup) + 1,
    Failed_Rank  = n() - rank(Failed) + 1
  ) %>%
  # convert and round the risk probabilities to percent
  mutate_at(vars(c(Attempt, Coup, Failed)), .funs = ~sprintf("%.1f", . * 100)) %>%
  select(Country, Attempt, Attempt_Rank, Coup, Coup_Rank, Failed, Failed_Rank) %>%
  arrange(Attempt_Rank)

datatable(fcasts, rownames = FALSE, 
          colnames = c("Attempt Rank" = 3, "Coup Rank" = 5, "Failed Rank" = 7),
          extensions = "Responsive",
          options = list(pageLength = 25))
```

You can download the forecasts [here (CSV)](https://raw.githubusercontent.com/andybega/forecaster2/master/forecasts/coup-2020.csv).
